---
title: "Class 2 Participation"
author: "Chip Galusha"
date: "1/20/2020"
output:
  md_document:
    variant: gfm
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(scipen = 999)

```

## Chip Galusha's Class One Participation Document
### Intro To ggPlot2

### Grammar of Graphics
ggplot2 implements the grammar of graphics - a coherent system for describing & building graphs. 

* Good grammar > 1st step for a good sentence   
* Good graphical grammar > 1st step in a well formed or correct graphic   

*ggplot2 builds up a graphic from multiple layers of data*   

**Components of the Layered Gramnmar**   
* data & aestedic mapping   
* geometric objects  
* scales
* facet specification
* statistical transformation   
* the coordinate system   

A layered grammar defines the components of a plot as:   
* A default dataset > mapping variables to aestedices
* One or more layer w/:
    - one geometric object   
    - one statistical transformation
    - one position adjustment   
    - and optionally a data set and aesdetic mapping
* one scale for each aestedic mapping   
* a coordinate system   
* a facet specification   

```{r}

#install.packages("gapminder")
library(gapminder)
library(tidyverse)
library(ggthemes)
library(wesanderson)

```

Let's get the data ready

```{r}
data("gapminder")

str(gapminder)
```

```{r}

ggplot(data = gapminder, mapping = aes(x = year, y = pop, color = continent)) +
  geom_point()

```

Aggregate by continent.

```{r}
gapminder %>% 
  group_by(year, continent) %>% 
  summarise(pop = sum(pop)) %>% 
ggplot(data = ., mapping = aes(x = year, y = pop, color = continent)) +
  geom_point() +
  scale_y_log10()

```

```{r}

gapminder %>% 
  filter(continent == "Asia") %>% 
ggplot(data = ., mapping = aes(x = year, y = gdpPercap, color = country)) +
  geom_point() +
  facet_wrap(~continent)
  

```

Try two numeric values.

```{r}

ggplot(data = gapminder, aes(x = gdpPercap, y = lifeExp, color = continent, size = pop)) +
  geom_point() +
  scale_x_log10() +
  stat_smooth(method = "lm")

```

Changing mapping for population.

```{r}
# this is useless but it shows how shape works
gapminder %>% 
  filter(continent == "Europe") %>%
  group_by(country) %>% 
  summarise(gdpPercap = mean(gdpPercap),
            lifeExp = mean(lifeExp)) %>% 
ggplot(data = ., mapping = aes(x = gdpPercap, y = lifeExp, shape = country)) +
  geom_point() +
  scale_x_log10() +
  stat_smooth(method = "lm")
```

Back to square one.

```{r}
ggplot(data = gapminder, mapping = aes(x = gdpPercap, y = lifeExp)) +
  geom_point(aes(color = continent), alpha = .4) +
  scale_x_log10() +
  stat_smooth(method = "lm", mapping = aes(color = continent))
```

```{r}
ggplot(data = gapminder, mapping = aes(x = gdpPercap, y = lifeExp)) +
  geom_point(alpha = .4) +
  scale_x_log10() +
  stat_smooth(method = "lm") +
  facet_wrap(~continent) +
  xlab("GDP per Capita") +
  ylab("Life Expectancy") +
  ggtitle("Life Expectancy By GDP per Capita", subtitle = " Faceted By Continent") 

```

Getting weird again

```{r}

gapminder %>% 
ggplot(data = ., mapping = aes(x = continent, y = lifeExp)) +
  stat_summary(fun.ymin = min,
               fun.ymax = max, 
               fun.y = median, 
               alpha = .9) +
  geom_point() +
  coord_flip() +
  theme_fivethirtyeight()

```


